---
title: "PartOne_Maimanah_MalariaDeath"
output: html_document
date: "2022-09-20"
---
#########################################

```{r warning=FALSE, message=FALSE}
library(knitr)

```


```{r}
url <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-13/malaria_deaths.csv"
data1 <- read.csv(url)
dim(data1)
names(data1)

url2 <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-13/malaria_deaths_age.csv"
data2 <- read.csv(url2)
dim(data2)
names(data2)

```





here is my Plots

```{r}
barplot(data1$Deaths...Malaria...Sex..Both...Age..Age.standardized..Rate...per.100.000.people., names.arg = data1$Year)
```
There is no trend for Global Malaria Deaths between 1990 ~ 2016, with unfiltered data.


```{r}
filter1 <- data1[data1$Year %in%c("1990"),  ]
names(filter1)
filter2 <- data1[data1$Entity %in% c("Uganda"),  ]
names(filter2)
filter3 <- data2[data2$entity %in% c("Uganda"),  ]
names(filter3)
filter4 <- filter3[filter3$year %in% c("1990"),  ]
names(filter4)

```
```{r}
barplot(filter1$Deaths...Malaria...Sex..Both...Age..Age.standardized..Rate...per.100.000.people., names.arg = filter1$Entity,
        main = "Year 1990")

```
Boxplot of Malaria Deaths in 1990 across the globe. Highest peak is the country of Uganda.


```{r}
barplot(filter2$Deaths...Malaria...Sex..Both...Age..Age.standardized..Rate...per.100.000.people., names.arg = filter2$Year,
        main = "Uganda")

```
Focusing on Uganda only, there is a downtrend of Malaria deaths from 1990 to 2016, with advancement of knowledge and medical facilities.

```{r}
barplot(filter3$deaths, names.arg = filter3$age_group,
        main = "Uganda Age Group Malaria 1990~2016")
```
Uganda Age-group of Malaria deaths, is extremely high for children under age of 5. Focus group to manage wil be under 5.


